---
title: 'Problem Set #1'
author: "Moretz_Brandon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---


```{r utility_functions, echo = F}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
	kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

getMode <- function(values) {

  tbl.vals <- data.frame(table(values))
  tbl.vals$values <- as.numeric(levels(tbl.vals$values))[tbl.vals$values]
  mode <- tbl.vals[which.max(tbl.vals$Freq),]

  names(mode)[1] <- "Value"

  mode
}

getSummary <- function(values) {
  mode <- getMode(values)

  data.table(
    Statistic = c("Mean",
          "Median",
          "Min",
          "Max",
          "Range",
          "Mode",
          "ModeFreq",
          "StdDev",
          "Variance",
          "Q1",
          "Q3",
          "P10"),

    Value = c(mean(values, na.rm = T),
        median(values, na.rm = T),
        min(values, na.rm = T),
        max(values, na.rm = T),
        max(values, na.rm = T) - min(values, na.rm = T),
        ifelse(mode$Freq != 1, mode$Value, NA),
        ifelse(mode$Freq != 1, mode$Freq, NA),
        sd(values, na.rm = T),
        var(values, na.rm = T),
        quantile(values, probs = c(0.25), na.rm = T),
        quantile(values, probs = c(0.75), na.rm = T),
        quantile(values, probs = c(0.10), na.rm = T))
  )
}

trimmed_mean <- function(vals, pct) {
  vals <- sort(vals)
  n <- length(vals)

  trim_vals <- floor(n * pct)

  trimmed <- head(vals, length(vals) - trim_vals)
  trimmed <- tail(trimmed, length(trimmed) - trim_vals)

  mean(trimmed)
}

winsorize <- function(vals, pct) {
  vals <- sort(vals)
  g <- floor(length(vals) * pct)

  bottom <- vals[g + 1]
  top <- vals[length(vals) - g]

  trimmed <- head(vals, length(vals) - g)
  trimmed <- tail(trimmed, length(trimmed) - g)

  c(rep(bottom, g), trimmed, rep(top, g))
}

```

## 1.) Extreme Outliers

Consider the values below. Using the boxplot rule, is the value 200 an extreme outlier ?

Values: 80, 121, 132, 145, 151, 119, 133, 134, 200, 195, 90, 121, 132, 123, 145, 151, 119, 133, 134, 151, 168,

```{r echo = T}

values <- c(80, 121, 132, 145, 151, 119, 133, 134, 200, 195, 90, 121, 132, 123, 145, 151, 119, 133, 134, 151, 168)

boxplot(values)

# boxplot rule
q1 <- quantile(values, .25)
q2 <- quantile(values, .75)

outlier_range <- q2 + 3 * (q2 - q1) # 3 * IQR for Extreme Outliers

pretty_vector( c( 200 > outlier_range) )

```

## 2.) Levels of Measurement

Determine which of the four levels of measurement (nominal, ordinal, interval, ratio) is most appropriate for calendar years.

+ Ratio 
+ Nominal 
+ __Interval__
+ Ordinal

__Answer:__ _Interval_

## 3.) Summary Stats

Here are some scores from a quiz.  Find the range for the given data.

```{r echo = T}

scores <- c(31, 47, 29, 31, 16, 48, 41, 50, 54, 37, 22)

pretty_vector( getSummary(scores)[c(5, 6),] )

```


## 4.) Standard Deviation

Find the sample standard deviation for the given sample data.  
Round your answer to one more decimal place than is present in the original data.

```{r echo = T}

sample <- c(18, 18, 17, 9, 15, 5, 10, 5, 15, 17, 7, 12, 12)

pretty_vector( getSummary(sample)[8] )

pretty_vector( round(sd(sample), 1), "Standard Deviation" )

```

## 5.) Median

+ Find the median for the given sample data.
+ The normal monthly precipitation(in inches) for August is listed for 20 different U.S. cities. Find the median of the data.

```{r echo = T}

monthly_precipitation <- c(3.6, 1.6, 2.4, 3.7, 4.1, 3.9, 1.0, 3.6, 4.2, 3.4, 3.7, 2.2, 1.5, 4.2, 3.5, 2.7, 0.4, 3.7, 2.0, 3.6)

pretty_vector( getSummary(monthly_precipitation)[2] )

```

## 6.) Eliminating Outlier

Multiple Choice 

Which of the following is not a reason for eliminating an identified outlier from a data set prior to further statistical analysis?

+ Faulty Distribution Assumption
+ Human error 
+ Intentional falsification 
+ __Sampling the wrong population__
+ Standardization failure 

## 7.) Chebyshev’s theorem

+ The heights of the adults in one town have a mean of 67.1 inches and a standard deviation of 3.5 inches.
+ What can you conclude from Chebyshev's theorem about the percentage of adults in the town whose heights are between 60.1 and 74.1 inches?  

_(Hint-study the section in Business Statistics that deals with this.)_

```{r echo = T}

mu <- 67.1
sd <- 3.5

pretty_vector(c("-2 sd" = mu - sd * 2, "+2 sd" = mu + sd * 2))

```


## 8.) Trimmed Mean

+ The data below consists of the heights (in inches) of 20 randomly selected women. 
+ Find the 20% trimmed mean of the data set.  Round to two decimal places. 

69, 68, 64, 61, 65, 64, 71, 67, 62, 63, 61, 64, 75, 67, 60, 59, 64, 69, 65, 72

```{r}

values <- c(69, 68, 64, 61, 65, 64, 71, 67, 62, 63, 61, 64, 75, 67, 60, 59, 64, 69, 65, 72)

# Base R
pretty_vector( round(mean(values, trim = .2), 2) )

# Custom Fun
pretty_vector( round(trimmed_mean(values, pct = .2), 2) )

```

## 9.) Group Standard Deviation

+ Find the standard deviation of the sample data summarized in the given frequency distribution.

```{r}

student.data <- data.table(
  Low = c(50, 60, 70, 80, 90),
  High = c(59, 69, 79, 89, 99),
  Students = c(5, 7, 9, 10, 9))

student.data[, Mid := (Low + High) / 2][, Freq := Mid * Students]
student.data[, Mf := Mid * Students][, MsF := (Mid ** 2) * Students]

mu <- sum(student.data$Mf) / sum(student.data$Students)

n <- sum(student.data$Students)
smsf <- sum(student.data$MsF)
smfs <- sum(student.data$Mf) ** 2 / n

var <- (smsf - smfs) / (n - 1)

pretty_vector( round(sqrt(var), 1) )

```

## 10.) Range & Variance

Find the range, variance, and standard deviation for each of the two samples, then compare the two sets of results.

When investigating times required for drive-through service, the following results (in seconds) were obtained.

```{r}

restaurantA <- c(120, 67, 89, 97, 124, 68, 72, 96)
restaurantB <- c(115, 126, 49, 56, 98, 76, 78, 95)

sum_cols <- c("Range", "Variance", "StdDev")

pretty_vector(getSummary(restaurantA)[which(Statistic %in% sum_cols)], "Restaurant A" )
pretty_vector(getSummary(restaurantB)[which(Statistic %in% sum_cols)], "Restaurant B" )

```
