---
title: 'Problem Set #1'
author: "Moretz_Brandon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---


```{r utility_functions}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

getMode <- function(values) {

  tbl.vals <- data.frame(table(values))
  tbl.vals$values <- as.numeric(levels(tbl.vals$values))[tbl.vals$values]
  mode <- tbl.vals[which.max(tbl.vals$Freq),]

  names(mode)[1] <- "Value"

  mode
}

getSummary <- function(values) {
  mode <- getMode(values)

  data.table(
    Statistic = c("Mean",
          "Median",
          "Min",
          "Max",
          "Range",
          "Mode",
          "ModeFreq",
          "StdDev",
          "Variance",
          "Q1",
          "Q3",
          "P10"),

    Value = c(mean(values, na.rm = T),
        median(values, na.rm = T),
        min(values, na.rm = T),
        max(values, na.rm = T),
        max(values, na.rm = T) - min(values, na.rm = T),
        ifelse(mode$Freq != 1, mode$Value, NA),
        ifelse(mode$Freq != 1, mode$Freq, NA),
        sd(values, na.rm = T),
        var(values, na.rm = T),
        quantile(values, probs = c(0.25), na.rm = T),
        quantile(values, probs = c(0.75), na.rm = T),
        quantile(values, probs = c(0.10), na.rm = T))
  )
}

trimmed_mean <- function(vals, pct) {
  vals <- sort(vals)
  n <- length(vals)

  trim_vals <- floor(n * pct)

  trimmed <- head(vals, length(vals) - trim_vals)
  trimmed <- tail(trimmed, length(trimmed) - trim_vals)

  mean(trimmed)
}

winsorize <- function(vals, pct) {
  vals <- sort(vals)
  g <- floor(length(vals) * pct)

  bottom <- vals[g + 1]
  top <- vals[length(vals) - g]

  trimmed <- head(vals, length(vals) - g)
  trimmed <- tail(trimmed, length(trimmed) - g)

  c(rep(bottom, g), trimmed, rep(top, g))
}

```

## 1.)

Consider the values below. Using the boxplot rule, is the value 200 an extreme outlier ?

Values: 80, 121, 132, 145, 151, 119, 133, 134, 200, 195, 90, 121, 132, 123, 145, 151, 119, 133, 134, 151, 168,

```{r echo = T}

values <- c(80, 121, 132, 145, 151, 119, 133, 134, 200, 195, 90, 121, 132, 123, 145, 151, 119, 133, 134, 151, 168)

boxplot(values)

# boxplot rule
q1 <- quantile(values, .25)
q2 <- quantile(values, .75)

outlier_range <- q2 + 3 * (q2 - q1) # 3 * IQR for Extreme Outliers

200 > outlier_range

```

## 2.)

Determine which of the four levels of measurement (nominal, ordinal, interval, ratio) is most appropriate for calendar years.

```{r echo = T}

print("Years <- Interval")

```

## 3.)

Here are some scores from a quiz.  Find the range for the given data.

```{r echo = T}

scores <- c(31, 47, 29, 31, 16, 48, 41, 50, 54, 37, 22)

getSummary(scores)[c(5, 6),]
summary(scores)

```


## 4.)

Find the sample standard deviation for the given sample data.  
Round your answer to one more decimal place than is present in the original data.

```{r echo = T}

sample <- c(18, 18, 17, 9, 15, 5, 10, 5, 15, 17, 7, 12, 12)

getSummary(sample)[8]
round(sd(sample), 1)

```

## 5.)

+ Find the median for the given sample data.
+ The normal monthly precipitation(in inches) for August is listed for 20 different U.S. cities. Find the median of the data.

```{r echo = T}

monthly_precipitation <- c(3.6, 1.6, 2.4, 3.7, 4.1, 3.9, 1.0, 3.6, 4.2, 3.4, 3.7, 2.2, 1.5, 4.2, 3.5, 2.7, 0.4, 3.7, 2.0, 3.6)
getSummary(monthly_precipitation)[2]

```

## 6.)

Multiple Choice 

Which of the following is not a reason for eliminating an identified outlier from a data set prior to further statistical analysis?

<br />
__Faulty Distribution Assumption__
<br />
Human error 
<br />
Intentional falsification 
<br />
Sampling the wrong population 
<br />
Standardization failure 

## 7.)

+ The heights of the adults in one town have a mean of 67.1 inches and a standard deviation of 3.5 inches.
+ What can you conclude from Chebyshev's theorem about the percentage of adults in the town whose heights are between 60.1 and 74.1 inches?  

_(Hint-study the section in Business Statistics that deals with this.)_

```{r echo = T}

mu <- 67.1
sd <- 3.5
mu + sd * 2
mu - sd * 2

```


## 8.)

+ The data below consists of the heights (in inches) of 20 randomly selected women. 
+ Find the 20% trimmed mean of the data set.  Round to two decimal places. 

69, 68, 64, 61, 65, 64, 71, 67, 62, 63, 61, 64, 75, 67, 60, 59, 64, 69, 65, 72

```{r}

values <- c(69, 68, 64, 61, 65, 64, 71, 67, 62, 63, 61, 64, 75, 67, 60, 59, 64, 69, 65, 72)

# Base R
round(mean(values, trim = .2), 2)

# Custom Fun
round(trimmed_mean(values, pct = .2), 2)

```

## 9.)

+ Find the standard deviation of the sample data summarized in the given frequency distribution.

```{r}

student.data <- data.table(
  Low = c(50, 60, 70, 80, 90),
  High = c(59, 69, 79, 89, 99),
  Students = c(5, 7, 9, 10, 9))

student.data[, Mid := (Low + High) / 2][, Freq := Mid * Students]
student.data[, Mf := Mid * Students][, MsF := (Mid ** 2) * Students]

mu <- sum(student.data$Mf) / sum(student.data$Students)

n <- sum(student.data$Students)
smsf <- sum(student.data$MsF)
smfs <- sum(student.data$Mf) ** 2 / n

var <- (smsf - smfs) / (n - 1)

round(sqrt(var), 1)

```

## 10.)

```{r}

restaurantA <- c(120, 67, 89, 97, 124, 68, 72, 96)
restaurantB <- c(115, 126, 49, 56, 98, 76, 78, 95)

sum_cols <- c("Range", "Variance", "StdDev")

getSummary(restaurantA)[which(Statistic %in% sum_cols)]
getSummary(restaurantB)[which(Statistic %in% sum_cols)]

```
