---
title: 'Problem Set #2'
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

## 1.)

Assume the outcomes from an experiment with a trials conform to a Poisson distribution with lambda equal to 1. 

__Determine the probability of obtaining an outcome greater than 3.__

_(Round to four decimal places.)_


```{r echo = T}

# method a, use lower.tail = F
prob <- ppois(3, lambda = 1, lower.tail = F)
round(prob, 4)

# method 2, use 1 - prob(3)
round(1 - ppois(3, lambda = 1), 4)

lam <- 1
x = seq( from = 0, to = 5, by = 1)

pdf = dpois(x, lambda = lam)
plot(x, pdf, type = "h", lwd = 3, col = "blue",
  main = paste("PDF of POIS(", lam,") with Approximating Normal Density"))
abline(h = 0, col = "green2")
curve(dnorm(x, 1, sqrt(1)), lwd = 2, col = "red", add = T)

```

## 2.)

__Use Bayes' theorem to find the indicated probability.__

5.8% of a population is infected with a certain disease. 

There is a test for the disease, however the test is not completely accurate. 

93.9% of those who have the disease test positive. 
However 4.1% of those who do not have the disease also test positive (false positives).
A person is randomly selected and tested for the disease.

> __What is the probability that the person has the disease given that the test result is positive?__


```{r echo = T}

d <- 0.058
dNot <- (1 - d)

dHaveTrue <- 0.939
dHaveFalse <- (1 - dHaveTrue)

dNotTrue <- 0.041 #False Positive
dNotFalse <- (1 - dNotTrue)

Q2 <- (d * dHaveTrue) / ((dNot * dNotTrue) + (d * dHaveTrue))
round(Q2, 3)

```

## 3.)

> Given the following probability distribution for the variable x, compute the expected value and variance of x using the distribution below.  

<table style="width: 516px;">
<tbody>
<tr>
<td style="width: 137px; text-align: center;">x</td>
<td style="width: 77px; text-align: center;">0</td>
<td style="width: 86px; text-align: center;">1</td>
<td style="width: 82px; text-align: center;">2</td>
<td style="width: 89px; text-align: center;">3</td>
</tr>
<tr>
<td style="width: 137px; text-align: center;">probability</td>
<td style="width: 77px; text-align: center;">0.779</td>
<td style="width: 86px; text-align: center;">0.195</td>
<td style="width: 82px; text-align: center;">0.024</td>
<td style="width: 89px; text-align: center;">0.002</td>
</tr>
</tbody>
</table>
<p>

_Round to two decimal places._

```{r echo = T}

# Basic Statistics, p49-50

vals <- c(0, 1, 2, 3)
probs <- c(0.779, 0.195, 0.024, 0.002)

ev <- sum( vals * probs )

round(ev, 2)

var <- sum((vals - ev) ** 2 * probs)
round(var, 2)

```


## 4.)

_Solve the problem._

Given the following sample data: __1.3, 2.2, 2.7, 3.1, 3.3, 3.7__, use quantile() in R with type = 7 to find the estimated 33rd percentile.

_Round to 2 decimal places._

```{r echo = T}

sample_data <- c(1.3, 2.2, 2.7, 3.1, 3.3, 3.7)

prob <- quantile(x = sample_data, probs = .33, type = 7)

round(prob, 2)

```

## 5.)

__Solve the problem.__

A study of the amount of time it takes a mechanic to rebuild the transmission for a 2005 Chevrolet Cavalier shows that 
the mean is 8.4 hours and the standard deviation is 1.8 hours. 

> If a random sample of 36 mechanics is selected, find the probability that their mean rebuild time exceeds 8.7 hours.  

_Assume the mean rebuild time has a normal distribution.  (Hint, interpolate in the tables or use pnorm().)_

```{r echo = T}

# Business Statistics, 7.2, p.219

ev <- 8.7
u <- 8.4
sd <- 1.8
n <- 36

z <- (ev - u) / (sd / sqrt(n))

round( pnorm( z, lower.tail = F ), 3)

```

## 6.)

> Use the normal distribution to estimate the probability of 50 successes for a 
binomial distribution with n = 67 and the probability of success p = 0.7.

_Round to four decimal places._

```{r echo = T}

# Business Statistics, Section 6.6, p.188
n <- 67
p = 0.7
q <- 1 - p

value <- 50

u <- n * p
sd <- sqrt(n * p * q)

stopifnot(u + 3 * sd < n)
stopifnot(u - 3 * sd < n)

z_1 <- (value - .5 - u) / sd
prob_1 <- pnorm(z_1, lower.tail = T)

z_2 <- (value + .5 - u) / sd
prob_2 <- pnorm(z_2, lower.tail = T)

round( prob_2 - prob_1, 4 )

```

## 7.)

_The given values are discrete (binomial outcomes)._

__Use the continuity correction and describe the region of the normal distribution that corresponds to the indicated probability.__

> The probability of less than 44 correct answers.

```{r echo = T}

# "Correcting for Continuity", Business Statistics, p.186
# The area to the left of 43.5

vals <- 1:100
probs <- dnorm(vals, mean = 50, 15 )
area_split <- 43.5

stopifnot(round(sum(probs)) == 1)

plot(vals, probs, type = "h", lwd = 3, col = ifelse(vals < area_split, "red", "blue"),
  main = "Probability of Values", xlab = "Value", ylab = "Probability")

abline(h = 0, col = "green2")

```


## 8.)

_Solve the problem. Round to the nearest tenth unless indicated otherwise._

> In one region, the September energy consumption levels for single-family homes are found to be 
normally distributed with a mean of 1050 kWh and a standard deviation of 218 kWh. Find P45 (45th percentile).

```{r}

Q8 <- qnorm(.45, mean = 1050, sd = 218, lower.tail = T)
round(Q8, 1)

```

## 9.)

_Find the indicated probability._

> In a taste test, five different customers are each presented with 3 different soft drinks
The same soft drinks are used with each customer, but presented in random order. 
If the selections were made by random guesses, find the probability that all five customers witnesses would pick the same soft drink as their favorite.  

_(There is more than one way the customers can agree.)_

```{r}

Q9 <- dbinom(x = 4, size = 4, prob = 1 / 3)
round(Q9, 5)

# or, 4 consecutive picks of the same drink (same as the first).
round( (1 / 3) ^ 4, 5)

```

## 10.)

_Find the indicated probability._

> As a prize for winning a contest, the contestant is blindfolded and allowed to draw 3 dollar bills one at a time out of an urn.
The urn contains forty $1 bills and ten $100 bills. 
The urn is churned before each selection so that the selection would be at random.
What is the probability that none of the $100 bills are selected? 

_(Round to four decimal places.)_

```{r}

Q10 <- dhyper( x = 0, m = 10, n = 40, k = 3 )
round(Q10, 4)

```
