---
title: 'Problem Set #3'
author: "Moretz_Brandon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    df_print: paged
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig, big.mark = "'") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
  pander::pander(vec)
}

normal_z <- function(pct, mean = 0, sd = 1) {
	qnorm(1 - ((1 - pct) / 2), mean, sd, lower.tail = T)
}

```


## 1.) Power Test

+ Assume a random sample of size n is from a normal population.
+ Assume a single sample t test is used to for hypothesis testing.
+ The null hypothesis is that the population mean is zero versus the alternative hypothesis that it is not zero.

If the sample size is decreased, and the Type I error rate is unchanged, then the Type II error rate will increase.

__Answer__

> False.
<br/><br/>
β varies inversley with sample size and α varies directly with sample size (as sample size increases, 
β decreases, α increases, power increases).

_Business Statistics, p.275._


## 2.) Critical Value

+ Assume the population has a normal distribution and the number of observations in a random sample is greater than fifty.

If a z test is used to test a null hypothesis about the population mean versus the alternative that the true mean is greater than 
that specified by the null hypothesis, what is the critical value for a one-tailed test?  

_Assume the type one-error rate is 0.05._

```{r echo = T}

pretty_vector( qnorm(.95), 3 )

```

## 3.) Confidence Interval

Use the degree of confidence and sample data to construct a confidence interval for the population proportion p.

n = 56, x = 30; 95% confidence

_(Use the procedure in Business Statistics Section 8.3.)_

```{r echo = T}

n <- 56
x <- 30
zValue <- normal_z(.95)
pHat <- x / n
moe <- zValue * sqrt((pHat * ((1 - pHat) / n)))
ciLower <- pHat - moe
ciUpper <- pHat + moe
pretty_vector(round(c(ciLower, pHat, ciUpper), 3))

```

## 4.) Sample Size

Use the given data to find the sample size required to estimate the population proportion.

Margin of error: 0.012; confidence level: 98%; p and q unknown

```{r echo = T}

z <- qnorm( 1 - ( 1 - 0.98) / 2)
moe <- 0.012
n = .25 * ( z / moe ) ** 2

pretty_vector(n)

```

## 5.) Distribution Characteristics

Which of the following statements is not true?

+ a.) The sampling distribution of a statistic is the probability distribution for that statistic based on all possible random samples from a population.
+ b.) A symmetric, heavy-tailed distribution cannot be detected using a boxplot and QQ chart.
+ c.) The t-statistic distribution is based upon the assumption of a random sample from a normal population.
+ d.) If the sample size is increased and the same test statistic and alpha level are used, the power of the test will increase.

__Answer:__

Choice __b.)__ is false. The QQ plot is used to determine wheather the data is normally distributed.

## 6.) Minimum Sample Size

Use the information to find the minimum sample size required to estimate an unknown population mean µ. 

Margin of error: $135, confidence level: 95%, σ = $500.

```{r echo = T}

moe <- 135
zValue <- qnorm(1 - 0.05 / 2)
sigma <- 500
n <- ((zValue ** 2 * sigma ** 2) / (moe ** 2))

pretty_vector(round( n ))

```

## 7.) Confidence Interval

_Use the information provided to determine a 95% confidence interval for the population variance._

+ A researcher was interested in the variability in service time (in hours) spent by mechanics fixing the same automotive problem. 
+ A random sample was taken resulting in a sample of size 20 from a substantial file of reported experience.
+ The summary statistics are as follows:  n = 20, sample mean = 13.8 hours, sample standard deviation = 3.9 hours.
+ Assume service time follows a normal distribution.  

__Round to two decimal places.__

```{r echo = T}

n <- 20
mu <- 13.8
std <- 3.9
conf.level = 0.95

df <- n - 1
chilower <- qchisq((1 - conf.level) / 2, df, lower.tail = TRUE)
chiupper <- qchisq((1 - conf.level) / 2, df, lower.tail = FALSE)
v <- std ** 2
conf.int <- c(df * v / chiupper, df * v / chilower)
pretty_vector(round(conf.int, 2))

```


## 8.) Confidence Interval

_Solve the problem._

+ A 99% confidence interval (in inches) for the mean height of a population is 65.7 < µ < 67.3.
+ This result is based on a sample of size 144. Construct the 95% confidence interval. 

_(Hint: you will first need to find the sample mean and sample standard deviation)._

```{r echo = T}

n <- 144

ci.l <- 65.7
ci.h <- 67.3

margin.of.error <- ( ci.h - ci.l ) / 2

mu_1 <- ci.l + margin.of.error
mu_2 <- ci.h - margin.of.error

stopifnot(mu_1 == mu_2)

mu <- mu_1

std <- margin.of.error / normal_z(.99) * sqrt(n)

moe <- normal_z(.95) * std / sqrt(n)

pretty_vector(c(mu - moe, mu + moe))

```

## 9.) Sampling Characteristics

Which of the following statements is __not true__ for sampling distributions?

+ A sampling distribution is necessary for making confidence statements about an unknown population parameter.
+ Depending on the population, it may not be possible to express the sampling distribution for a statistic in closed form mathematically.
+ A sampling distribution depends on the nature of the population being sampled.
+ When sampling at random from a normal population, the sampling distribution for the sample average is a normal distribution.
+ None of the above.

> __Answer:__

__None of the above.__

## 10.) P-Value

+ Assume normality and use the information given to find the p-value.

Based on the p-value estimated, determine if the null hypothesis should be rejected at a 0.1 significance level.  Select the correct answer if the 
test statistic in a two-tailed test is z= -1.63.  

Follow the procedure shown in Business Statistics.

__Business Statistics p.279__

When conducting two-tailed tests, recall that we  split alpha to determine the critical value of the test statistic. For a
two-tailed test, the p-value can be compared to α / 2 to reach a statistical conclusion.

If the p-value is that than α/2, the decision is to reject the null hypothesis.

```{r echo = T}

p <- pnorm(-1.63)
sig <- .1 / 2 # must divide by 2 in order to perform the two-tailed test.
pretty_vector(round( p, 3 ))
pretty_vector(ifelse(p < ( 0.1 / 2 ), "Reject the Null", "Fail to Reject the Null"))

```
