---
title: 'Week 5-6 Reading'
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```


```{r}
#rm(list = ls()) # Clear Workspace
```

# _Problem Set 3_

## 1.)

+ Assume a random sample of size n is from a normal population.
+ Assume a single sample t test is used to for hypothesis testing.
+ The null hypothesis is that the population mean is zero versus the alternative hypothesis that it is not zero.

If the sample size is decreased, and the Type I error rate is unchanged, then the Type II error rate will increase.    

```{r echo = T}

```

## 2.)

+ Assume the population has a normal distribution and the number of observations in a random sample is greater than fifty.

If a z test is used to test a null hypothesis about the population mean versus the alternative that the true mean is greater than that specified by the null hypothesis, what is the critical value for a one-tailed test?  

_Assume the type one-error rate is 0.05._

```{r echo = T}

```

## 3.)

Use the degree of confidence and sample data to construct a confidence interval for the population proportion p.

n = 56, x = 30; 95% confidence

_(Use the procedure in Business Statistics Section 8.3.)_

```{r echo = T}

qnorm(.95)

```

## 4.)

Use the given data to find the sample size required to estimate the population proportion.

Margin of error: 0.012; confidence level: 98%; p and q unknown

```{r echo = T}

```

## 5.)

```{r echo = T}

```

## 6.)

Use the information to find the minimum sample size required to estimate an unknown population mean µ. 

Margin of error: $135, confidence level: 95%, σ = $500.

```{r echo = T}

```

## 7.)

_Use the information provided to determine a 95% confidence interval for the population variance._

+ A researcher was interested in the variability in service time (in hours) spent by mechanics fixing the same automotive problem. 
+ A random sample was taken resulting in a sample of size 20 from a substantial file of reported experience.
+ The summary statistics are as follows:  n = 20, sample mean = 13.8 hours, sample standard deviation = 3.9 hours.
+ Assume service time follows a normal distribution.  

__Round to two decimal places.__

```{r echo = T}

```


## 8.)

_Solve the problem._

+ A 99% confidence interval (in inches) for the mean height of a population is 65.7 < µ < 67.3.
+ This result is based on a sample of size 144. Construct the 95% confidence interval. 

_(Hint: you will first need to find the sample mean and sample standard deviation)._

```{r echo = T}

```

## 9.)

```{r echo = T}

```

## 10.)

+ Assume normality and use the information given to find the p-value.

Based on the p-value estimated, determine if the null hypothesis should be rejected at a 0.1 significance level.  Select the correct answer if the test statistic in a two-tailed test is z= -1.63.  Follow the procedure shown in Business Statistics.

```{r echo = T}

```
