---
title: 'Week 5-6 Reading'
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```

```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

```

```{r utility}

normal_z <- function(pct, mean = 0, sd = 1) {
    qnorm( 1 - (( 1 - pct ) / 2), mean, sd, lower.tail = T)
}

```

# _Problem Set 3_

## 1.)

+ Assume a random sample of size n is from a normal population.
+ Assume a single sample t test is used to for hypothesis testing.
+ The null hypothesis is that the population mean is zero versus the alternative hypothesis that it is not zero.

If the sample size is decreased, and the Type I error rate is unchanged, then the Type II error rate will increase.

```{r echo = T}

# This is addressed in the sync session, page #2.

```

## 2.)

+ Assume the population has a normal distribution and the number of observations in a random sample is greater than fifty.

If a z test is used to test a null hypothesis about the population mean versus the alternative that the true mean is greater than that specified by the null hypothesis, what is the critical value for a one-tailed test?  

_Assume the type one-error rate is 0.05._

```{r echo = T}

```

## 3.)

Use the degree of confidence and sample data to construct a confidence interval for the population proportion p.

n = 56, x = 30; 95% confidence

_(Use the procedure in Business Statistics Section 8.3.)_

```{r echo = T}


```

## 4.)

Use the given data to find the sample size required to estimate the population proportion.

Margin of error: 0.012; confidence level: 98%; p and q unknown



```{r echo = T}

```

## 5.)

```{r echo = T}

```

## 6.)

Use the information to find the minimum sample size required to estimate an unknown population mean µ. 

Margin of error: $135, confidence level: 95%, σ = $500.

```{r echo = T}

```

## 7.)

_Use the information provided to determine a 95% confidence interval for the population variance._

+ A researcher was interested in the variability in service time (in hours) spent by mechanics fixing the same automotive problem. 
+ A random sample was taken resulting in a sample of size 20 from a substantial file of reported experience.
+ The summary statistics are as follows:  n = 20, sample mean = 13.8 hours, sample standard deviation = 3.9 hours.
+ Assume service time follows a normal distribution.  

__Round to two decimal places.__

```{r echo = T}

ci.pct = .95

n <- 20
mu <- 13.8
std <- 3.9

z <- normal_z(ci.pct)

margin.of.error <- z * std / sqrt(n)

c(mu - margin.of.error, mu + margin.of.error)

```


## 8.)

_Solve the problem._

+ A 99% confidence interval (in inches) for the mean height of a population is 65.7 < µ < 67.3.
+ This result is based on a sample of size 144. Construct the 95% confidence interval. 

_(Hint: you will first need to find the sample mean and sample standard deviation)._

```{r echo = T}

n <- 144

ci.l <- 65.7
ci.h <- 67.3

margin.of.error <- ( ci.h - ci.l ) / 2

mu_1 <- ci.l + margin.of.error
mu_2 <- ci.h + margin.of.error

stopifnot(mu1 == mu2)

mu <- mu1

std <- margin.of.error / normal_z(.99) * sqrt(n)

moe <- normal_z(.95) * std / sqrt(n)

round( c(mu - moe, mu + moe), 1 )

```

## 9.)

Which of the following statements is __not true__ for sampling distributions?

+ A sampling distribution is necessary for making confidence statements about an unknown population parameter.
+ Depending on the population, it may not be possible to express the sampling distribution for a statistic in closed form mathematically.
+ A sampling distribution depends on the nature of the population being sampled.
+ When sampling at random from a normal population, the sampling distribution for the sample average is a normal distribution.
+ None of the above.

> __Answer:__


## 10.)

+ Assume normality and use the information given to find the p-value.

Based on the p-value estimated, determine if the null hypothesis should be rejected at a 0.1 significance level.  Select the correct answer if the test statistic in a two-tailed test is z= -1.63.  Follow the procedure shown in Business Statistics.

```{r echo = T}

```
