---
title: "Lesson 09: Statistical Inference Hypothesis Testing for Single Populations"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```


```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
		kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
	data.raw <- fread(paste0(data.path, file_name),
	header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

	data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

	return(data)
}

#base.dir <- "E:/GitHub/R-Playground"
base.dir <- "C:/Projects/R/Playground"

data.path <- paste0(base.dir, "/Classes/MSDS401/Exercises/09_statistical_inference_hypothesis_testing_for_single_poulations/data/")

```


```{r data_load, include = F}

hotdogs <- loadDataFile("hot_dogs.csv")

```

# Questions

Variable Names:
+ 1. Type = Type of hotdog (beef, meat, or poultry)
+ 2. Calories = Calories per hot dog
+ 3. Sodium = Milligrams of sodium per hot dog

## 1.)

Use hot_dogs.csv data and hypothesis tests to determine which type of hot dog has average calories
less than 140 with 95% confidence. Present boxplots of calories by type of hot dog.

```{r question1}

```

## 2.)

Using hot_dogs.csv data and hypothesis tests at the 95% confidence level, determine which type of hot
dog has an average Sodium level different from 425 milligrams.

```{r question2}

```

## 3.)

Using hot_dogs.csv data and hypothesis tests, determine if the variance in Sodium values for beef hot
dogs is different from 6000 with 95% confidence.

```{r question3}

```

## 4.)

Assume a random sample of size 100 is drawn from a normal distribution for which the mean and
variance are unknown. Assume the sample mean is 50 and the standard deviation of the sample is 2.
Test the hypothesis that the true mean is 56, and also test the hypothesis the true mean is 40. Report
p-values and comment on the results.

```{r question4}

```

## 5.)

A coin is flipped 100 times. If it is unbiased the probability of a heads should equal the probability of a
tails. At the 95% confidence level, test the null hypothesis the coin is unbiased versus the alternative
that it is biased if 43 heads are obtained. Test the same hypothesis if 63 heads are obtained. Use
one-sided hypothesis tests.

```{r question5}

```

## 6.)

salaries.csv contains data derived from a November 8, 1993 article in Forbes titled “America’s Best
Small Companies”. The file gives the CEO age and salary for 60 small business firms. Use these data to
test the hypothesis at 95% confidence that at least 50% of the CEOs are 45 years old or older. Also
test the hypothesis at 95% confidence that at least 50% of the CEOs earn less than $500,000 per year.
Use one-sided hypothesis tests.

```{r question6}

```
