---
title: 'Data Analysis Assignment #1 (50 points total)'
author: "Moretz_Brandon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```

# _Abalones Data_

```{r analysis_setup, message = FALSE, warning = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr)
library(RColorBrewer)

theme_set(theme_fivethirtyeight())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
             axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             legend.position = "top", legend.title = element_blank())

#base.dir <- "E:/GitHub/R-Playground"
base.dir <- "C:/Projects/R/Playground"

data.path <- paste0(base.dir, "/Classes/MSDS401/Assignments/DataAnalysis_01/data/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
  data.raw <- fread(paste0(data.path, file_name),
    header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

  data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

  return(data)
}

# b) Use read.csv() to read the abalones.csv into R, assigning the data frame to "mydata."

mydata <- loadDataFile("abalones.csv") # this is equivilant to read.csv, however it returns a data.table.

# c) Use the str() function to verify the structure of "mydata."
glimpse(mydata)

# You should have 1036 observations.
stopifnot(nrow(mydata) == 1036) # verify observations.

# of eight variables.
stopifnot(ncol(mydata) == 8) # verify variables.

# d) Define two new variables, VOLUME and RATIO. 

# Use the following statements to define VOLUME and
# RATIO as variables appended to the data frame "mydata."

mydata$Volume <- mydata$Length * mydata$Diam * mydata$Height
mydata$Ratio <- mydata$Shuck / mydata$Volume

# Fix Sex Factor
mydata$Sex <- factor(x = mydata$Sex, labels = c("Infant", "Female", "Male"))

# Add row numbers to be used in sampling later.
mydata[, Num := .I]

```

-----

## 1.) Summary

### a.)
__(2 points)__

+ Use _summary()_ to obtain and present descriptive statistics from __mydata__.

```{r Part_1a_summary}

kable(summary(mydata), caption = "Summary")

```

+ Use _table()_ to present a frequency table using __CLASS__ and __RINGS__. _There should be 115 cells in the table you present._

```{r Part_1a_ft, eval = TRUE, echo = FALSE}

freq_tbl <- with(mydata, table(Class, Rings))

stopifnot(nrow(freq_tbl) * ncol(freq_tbl) == 115)

kable(addmargins(freq_tbl), caption = "Frequency Table: Class ~ Rings")

```

**Question (1 point):  Briefly discuss the variable types and distributional implications such as potential skewness and outliers.**

__Answer:__ 

### b.)
__(1 point)__ 

+ Generate a table of counts using __SEX__ and __CLASS__.

+ _Add margins to this table (Hint: There should be 15 cells in this table plus the marginal totals_.

+ Apply *table()* first, then pass the table object to *addmargins()* (Kabacoff Section 7.2 pages 144-147)).

+ Lastly, present a barplot of these data; ignoring the marginal totals.  

```{r Part_1b}

mytable <- with(mydata, table(Sex, Class))
kable(addmargins(mytable), caption = "Sex ~ Class")

# Side-by-side Stacked
ggplot(mydata, aes( Sex, fill = Sex)) +
	geom_bar() +
	facet_grid(~Class) +
	labs(title = "Comparison Abalone Sex Frequencies", x = "Age Class", y = "Frequency",
	    caption = "MSDS 401: Data Analysis 1, Q1_b") +
		theme(axis.text.x = element_text(size = 8))

```

**Question (2 points):  Discuss the sex distribution of abalones.  What stands out about the distribution of abalones by CLASS?**

__Answer:__

>

### c.)

__(1 point)__

+ Select a simple random sample of 200 observations from "__mydata__" and identify this sample as "__work__." 
+ Use *set.seed(123)* prior to drawing this sample. __Do not change the number 123__.
+ Note that *sample()* "takes a sample of the specified size from the elements of x." We cannot sample directly from "mydata." 
    Instead, we need to sample from the integers, 1 to 1036, representing the rows of "mydata." 
    Then, select those rows from the data frame (Kabacoff Section 4.10.5 page 87). 

__Using "work", construct a scatterplot matrix of variables 2-6 with *plot(work[, 2:6])*__
 
_(these are the continuous variables excluding VOLUME and RATIO). The sample "work" will not be used in the remainder of the assignment._

```{r Part_1c}

set.seed(123)

data.sample <- sample(x = 1:nrow(mydata), size = 200)

work <- mydata[Num %in% data.sample, .(Length, Diam, Height, Whole, Shuck)]

ggpairs(work) +
	labs(title = "Scatterplot Matrix",
		caption = "MSDS 401: Data Analysis 1, Q1.c") +
        theme(axis.text.x = element_text(size = 8), axis.text.y = element_text(size = 8))

```

-----

## 2.)

### a.) 

__(1 point)__

Use "mydata" to plot __WHOLE__ versus __VOLUME__. Color code data points by __CLASS__.

```{r Part_2a}

ggplot(mydata, aes(Volume, Whole, color = Class)) +
    geom_jitter() + # jitter to help with the overlap in the lower left near the origin
    labs(title = "Whole weight, in grams | Volume cm³",
        x = "Volume", y = "Whole Weight",
        caption = "MSDS 401: Data Analysis 1, Q2.a")

```


### b.) 

__(2 points)__

+ Use "mydata" to plot SHUCK versus WHOLE with WHOLE on the horizontal axis. 
+ Color code data points by CLASS. 
+ As an aid to interpretation, determine the maximum value of the ratio of SHUCK to WHOLE.
+ Add to the chart a straight line with zero intercept using this maximum value as the slope of the line. 
+ If you are using the 'base R' *plot()* function, you may use *abline()* to add this line to the plot. Use *help(abline)* in R to determine the coding for the 
    slope and intercept arguments in the functions. 
    If you are using ggplot2 for visualizations, *geom_abline()* should be used.

```{r Part_2b}

slope <- first(mydata[, .(Y = Shuck / Whole)][ order(-Y)]$Y)

ggplot(mydata, aes(x = Whole, y = Shuck, color = Class)) +
    geom_jitter() +
    geom_abline(slope = slope, intercept = 0, linetype = "dotdash", size = 1.2, color = "steelblue4", alpha = .7) +
    labs(title = "Shuck weight, in grams | Whole weight, in grams",
        x = "Whole Weight", y = "Shuck Weight",
        caption = "MSDS 401: Data Analysis 1, Q2.a")

```

**Question (2 points):  How does the variability in this plot differ from the plot in (a)?  
Compare the two displays.  Keep in mind that SHUCK is a part of WHOLE. 
Consider the location of the different age classes.**

__Answer:__

> 

-----

## 3.)

### a.) 
(2 points) 

+ Use "mydata" to create a multi-figured plot with histograms, boxplots and Q-Q plots of RATIO differentiated by sex. 
+ This can be done using *par(mfrow = c(3,3))* and base R or *grid.arrange()* and ggplot2.
+ The first row would show the histograms, the second row the boxplots and the third row the Q-Q plots. 
+ Be sure these displays are legible.  

```{r Part_3a_hist}

# I broke these 3 out into seperate R chuncks to give natural borders to the consecutive plots.

ggplot(mydata, aes(x = Ratio, fill = ..count..)) +
	geom_histogram(bins = 15) +
	facet_wrap(aes(Sex)) +
    labs( main = "Sex ~ Ratio Distribution" ) +
	theme(axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"))

```


```{r Part_3a_box}

ggplot(mydata, aes(x = Sex, y = Ratio)) +
	geom_boxplot(outlier.colour = "red", fill = "#20D4EA", outlier.shape = 1)


```


```{r Part_3a_qq}

ggplot(mydata, aes(sample = Ratio)) +
	stat_qq(color = "steelblue4") +
	stat_qq_line(lwd = .8, alpha = .8) +
	facet_wrap(aes(Sex))

```

**Question (2 points): Compare the displays.  How do the distributions compare to normality? Take into account the criteria discussed in the sync sessions to evaluate non-normality.**

__Answer:__

***(Enter your answer here.)***

### b.)
(2 points) 

+ Use the boxplots to identify RATIO outliers (mild and extreme both) for each sex. 
+ Present the abalones with these outlying RATIO values along with their associated variables in "mydata"

_(Hint:display the observations by passing a data frame to the kable() function)._

```{r Part_3b}


```

**Question (2 points):  What are your observations regarding the results in (3)(b)?**

__Answer:__

***(Enter your answer here.)***

-----

(4)(a) (3 points) With "mydata," display side-by-side boxplots for VOLUME and WHOLE, each differentiated by CLASS (Davies Section 14.3.2). There should be five boxes for VOLUME and five for WHOLE. Also, display side-by-side scatterplots:  VOLUME and WHOLE versus RINGS. Present these four figures in one graphic:  the boxplots in one row and the scatterplots in a second row. Base R or ggplot2 may be used.


```{r Part_4a}


```

**Question (5 points) How well do you think these variables would perform as predictors of age?  Explain.**

***Answer: (Enter your answer here.)***

-----

(5)(a) (2 points) Use *aggregate()* with "mydata" to compute the mean values of VOLUME, SHUCK and RATIO for each combination of SEX and CLASS. Then, using *matrix()*, create matrices of the mean values. Using the "dimnames" argument within *matrix()* or the *rownames()* and *colnames()* functions on the matrices, label the rows by SEX and columns by CLASS. Present the three matrices (Kabacoff Section 5.6.2, p. 110-111). The *kable()* function is useful for this purpose.  You do not need to be concerned with the number of digits presented.

```{r Part_5a}


```

(5)(b) (3 points) Present three graphs. Each graph should include three lines, one for each sex. The first should show mean RATIO versus CLASS; the second, mean VOLUME versus CLASS; the third, mean SHUCK versus CLASS. This may be done with the 'base R' *interaction.plot()* function or with ggplot2 using *grid.arrange()*.

```{r Part_5b, fig.width = 9}


```

**Question (2 points):  What questions do these plots raise?  Consider aging and sex differences.**

***Answer: (Enter your answer here.)***

5(c) (3 points) Present four boxplots using *par(mfrow = c(2, 2)* or *grid.arrange()*. The first line should show VOLUME by RINGS for the infants and, separately, for the adult; factor levels "M" and "F," combined. The second line should show WHOLE by RINGS for the infants and, separately, for the adults. Since the data are sparse beyond 15 rings, limit the displays to less than 16 rings. One way to accomplish this is to generate a new data set using subset() to select RINGS < 16.  Use ylim = c(0, 1100) for VOLUME and ylim = c(0, 400) for WHOLE.  If you wish to reorder the displays for presentation purposes or use ggplot2 go ahead.

```{r Part_5c}


```

**Question (2 points):  What do these displays suggest about abalone growth?  Also, compare the infant and adult displays.  What differences stand out?**

***Answer: (Enter your answer here.)***

-----

**Conclusions**

Please respond to each of the following questions (10 points total):

**Question 1) (5 points)	Based solely on these data, what are plausible statistical reasons that explain the failure of the original study? Consider to what extent physical measurements may be used for age prediction.**

***Answer: (Enter your answer here.)***

**Question 2) (3 points) Do not refer to the abalone data or study.  If you were presented with an overall histogram and summary statistics from a sample of some population or phenomenon and no other information, what questions might you ask before accepting them as representative of the sampled population or phenomenon?**

***Answer: (Enter your answer here.)***

**Question 3)  (2 points)	Do not refer to the abalone data or study.  What do you see as difficulties analyzing data derived from observational studies? Can causality be determined?  What might be learned from such studies?**

***Answer: (Enter your answer here.)***
