---
title: "Using R for Statistics"
author: "Moretz, Brandon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

## Probability Distributions in R

### 1.)

> Suppose that a fair die is rolled 10 times. What is the probability of throwing exactly two sixes?

```{r}
    dbinom(2, 10, 1/6)
```

### 2.)
> The number of lobster ordered in a restaurant on a given day is known to follow a Poisson distribution with a mean of 20.
What is the probability that exactly eighteen lobsters will be ordered tomorrow?

```{r}
    dpois(18, 20)
```

### 3.)

> To find the value of the pdf at 'x = 2.5' for a normal distribution with a mean of 5 and a
standard deviation of 2, use:

```{r}
    dnorm(2.5, mean = 5, sd = 2)
```

## Finding Probabilities

__Normal Distribution__

> To find the probability that a randomly selected value will be less than or eq1ual to 2.5 on the 

```{r}
    pnorm(2.5)
```
> Find a probability for a nonstandard normal distribution, add mean and sd params:

```{r}
    pnorm( 6, mean = 5, sd = 2)
```

> To find the complementary probability that the value will be _greater_ than 6, set lower.tail = F.

```{r}
    pnorm(6, 5, 2, lower.tail = F)
```

__which is equivalent to__

```{r}
    1 - pnorm(6, 5, 2)
```

### The Normal Distribution

> Suppose the height of men in the United Kingdom is known to be normally distributed with a mean of 177 centimeters a
and a standard deviation of 10 centimeters. If you were to select a man from the United Kiungdom population at random,
what is the probability that he would be more than 200 centimeters tall?

```{r}
    pnorm(200, mean = 177, sd = 10, lower.tail = F)
```

> What is the probability that he would be less than 150 centimeters tall?

```{r}
    pnorm(150, mean = 177, sd = 10)
```

### Binomial Distribution

> If you were to roll a fair six-sided die, 100 times, what is the probability of rolling a six no more than 10 times?

_the number of sixes in 100 dice roll rollows a binomial distribution, so you can answer the question with pbinom._

+ The reasoning for this is that we are interested in a binary event, is the die a 6? __Yes | no__
+ How many rolls must equal 6, out of how many total rolls?
+ What's the chance of rolling a 6? 1/6

```{r}
    pbinom(10, 100, 1/6)
```

> What is the probability of rolling a six more than 20 times?

```{r}
    pbinom(20, 100, 1/6, lower.tail = F)
```

### Exponential Distribution

Malfunctions in a particular type of electronic device are known to follow an exponential distribution with a mean time of 24
months until the device malfunctions. What is the probability that a randomly selected device will malfunction within the
first 6 months?

```{r}
    pexp(6, 1/24)
```

> What is the probability that a randomly selected device will last more than 5 years (60 months) without malfunction?

```{r}
    pexp(60, 1/24, lower.tail = F)
```

## Quantiles

> To find quantiles of the Normal Distribution, we can use the qnorm function:

```{r}
    qnorm(.95)
```

> For nonstandard normal distributions, use the mean and sd arguments to specifiy the parameters for the distribution.

```{r}
    qnorm(.95, mean = 5, sd = 2)
```

> To find the value _above_ which 95% of the population falls, set the lower.tail = F.

```{r}
    qnorm( .95, mean = 5, sd = 2, lower.tail = F)
```

> To find the value below which 0.5 percent of the population falls, use the command:

```{r}
    qnorm(0.005, 195, 17)
```


```{r}
    qnorm(0.005, 195, 17, lower.tail = F)
```

> To find the length of time within with 40% of devices will have malfunctioned, use the command:

```{r}
    qexp(0.4, 1/24)
```
_40% of devices will fail within 12.3 months._

> The number of lobsters ordered in a resturant on a given day is known to follow a Poisson distribution with
a mean of 20. If the manager wants to be able to satisfy all requests for lobster on at least 80% of days,
how many lobsters should they order each day?


```{r}
    qpois(.8, 20)
```
_by ordering 24 lobsters per day, the restaurant will be able to satisfy all requesat for lobster on at least 80% of days._

## Random Numbers

> To generate 100 random numbers from a standard normal distribution:

```{r}
    rnorm(100)
```

> To generate a simple random sample from a range of numbers, you can use the sample function. 

For example, to select 20 random numbers between 1 and 100 without replacement:

```{r}
    sample(1:100, 20, replace = T)
```

### Random Normal

```{r}
    rnorm(3, 195, 17)
```

### Binomial

```{r}
    rbinom(1, 10, 1/6)
```

### Poisson

```{r}
    rpois(7, 20)
```

### Exponental

```{r}
    rexp(10, 1/24)
```

## Curves

```{r}
    curve(dnorm(x))
```
