---
title: "Data Analysis #2"
author: "Moretz, Brandon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```

# __Abalones__

__75 points total__

_Raw Data_

```{r analysis_setup1, message = FALSE, warning = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

# Perform the following steps to start the assignment.
 
# 1) Load/attach the following packages via library():  flux, ggplot2, gridExtra, moments, rockchalk, car.
# NOTE:  packages must be installed via install.packages() before they can be loaded.

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(moments, quietly = TRUE, warn.conflicts = FALSE)
library(flux, quietly = TRUE, warn.conflicts = FALSE)
library(rockchalk, quietly = TRUE, warn.conflicts = FALSE)
library(car, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

pretty_kable <- function(data, title, dig = 2) {
	kable(data, caption = title, digits = dig, big.mark = "'") %>%
		kable_styling(bootstrap_options = c("striped", "hover"))
}

pretty_vector <- function(vec, label = "") {
	pander::pander(vec)
}

#base.dir <- "E:/GitHub/R-Playground"
base.dir <- "C:/Projects/R/Playground"

data.path <- paste0(base.dir, "/Classes/MSDS401/Assignments/DataAnalysis_02/data/")

# simple replacement for read.csv that returns a data.table.
loadDataFile <- function(file_name) {
	data.raw <- fread(paste0(data.path, file_name),
	header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", ""))

	data <- setNames(data.raw, tools::toTitleCase(tolower(names(data.raw))))

	return(data)
}

# 2) Use the "mydata.csv" file from Assignment #1 or use the file posted on the course site.  Reading
# the files into R will require sep = "" or sep = " " to format data properly.  Use str() to check file
# structure.

mydata <- loadDataFile("mydata.csv")

str(mydata)

```

-----

## 1.) Ratio

### a.) Kurtosis

__(1 point)__

+ Form a histogram and QQ plot using __RATIO__. 
+ Calculate skewness and kurtosis using '__rockchalk__.' 

Be aware that with 'rockchalk', the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package.

```{r Part_1a}

stopifnot(!is.na(mydata$Ratio))

ggplot(mydata, aes(Ratio, fill = ..count..)) +
   geom_histogram(breaks = pretty(mydata$Ratio)) +
   labs(title = "Histogram of Ratio", caption = "MSDS 401: Data Analysis #2, Q1.a") +
   theme(legend.text = element_blank())

ggplot(mydata, aes(sample = Ratio)) +
   stat_qq(color = "steelblue3") +
   stat_qq_line(lwd = 1, linetype = 12) +
   labs(title = "QQ of Ratio", caption = "MSDS 401: Data Analysis #2, Q1.a")

```


```{r Part_1a_kurtosis}
pretty_vector(c(Kurtosis = rockchalk::kurtosis(mydata$Ratio)))
```

### b.) Log Scale

__(2 points)__

+ Tranform RATIO using *log10()* to create L_RATIO (Kabacoff Section 8.5.2, p. 199-200). 
+ Form a histogram and QQ plot using __L_RATIO__. 
+ Calculate the skewness and kurtosis. 
+ Create a boxplot of __L_RATIO__ differentiated by __CLASS__.

```{r Part_1b}

mydata$L_Ratio <- log10(mydata$Ratio)

stopifnot(!is.na(mydata$L_Ratio))

ggplot(mydata, aes(L_Ratio, fill = ..count..)) +
   geom_histogram(breaks = pretty(mydata$L_Ratio)) +
   labs(title = "Histogram of L_Ratio", caption = "MSDS 401: Data Analysis #2, Q1.b") +
   theme(legend.text = element_blank())

ggplot(mydata, aes(sample = L_Ratio)) +
   stat_qq(color = "steelblue3") +
   stat_qq_line(lwd = 1, linetype = 12) +
   labs(title = "QQ of L_Ratio", caption = "MSDS 401: Data Analysis #2, Q1.b")

```

```{r Part_1b_skewkurt}

pretty_vector(c( Kurtosis = rockchalk::kurtosis(mydata$L_Ratio), Skewness = rockchalk::skewness(mydata$L_Ratio) ))

```


```{r Part_1b_lratiobyclass}

ggplot(mydata, aes(x = Class, y = L_Ratio, group = Class)) +
   geom_boxplot(outlier.colour = "red", fill = "steelblue3", outlier.shape = 1)

```

### c.) Bartlett's Test
__(1 point)__

+ Test the homogeneity of variance across classes using *bartlett.test()* 

_(Kabacoff Section 9.2.2, p. 222)._

```{r Part_1c}

bartless.l_ratio <- bartlett.test( L_Ratio ~ Class, data = mydata)

qqPlot(lm(L_Ratio ~ Class, data = mydata),
	simulate = T, main = "QQ-Plot L_Ratio", labels = F)

bartless.l_ratio

pretty_vector(c(P_Value = ifelse(bartless.l_ratio$p.value < 0.05, "Reject Null", "Cannot Reject Null")))

bartlett.ratio <- bartlett.test( Ratio ~ Class, data = mydata)

qqPlot(lm(Ratio ~ Class, data = mydata),
	simulate = T, main = "QQ-Plot Ratio", labels = F)

bartless.ratio

pretty_vector(c(P_Value = ifelse(bartlett.ratio$p.value < 0.05, "Reject Null", "Cannot Reject Null")))
```

**Question (2 points):  Based on steps 1.a, 1.b and 1.c, which variable RATIO or L_RATIO exhibits better conformance to a normal distribution with homogeneous variances across age classes?  Why?** 

__Answer:__

__L_Ratio__ exhibits better conformance to normality across classes due to the QQ-Plots falling closer in-line with the expected quantiles, lower Kurtosis, as well as having a less skewed 
distribution as observed by both the skewness measure and the histograms of the respective variables. Finally, at 0.05 significance level, we cannot reject the null hypothesis that
L_Ratio comes from a normal distribution according to Bartless's Test for Equality of Variances, which we must do for Ratio as if falls well short of significance.

-----

### 2.) Analysis of Variance

### a.) 

__(2 points)__

+ Perform an analysis of variance with *aov()* on L_RATIO using CLASS and SEX as the independent variables _(Kabacoff chapter 9, p. 212-229)_. 
+ Assume equal variances. 
+ Perform two analyses. 
+ First, fit a model _with_ the interaction term __CLASS:SEX__. 
+ Then, fit a model _without_ __CLASS:SEX__. 
+ Use *summary()* to obtain the analysis of variance tables (Kabacoff chapter 9, p. 227).

```{r Part_2a}

anova_1 <- aov(L_Ratio ~ Class * Sex, data = mydata)
summary(anova_1)

anova_2 <- aov(L_Ratio ~ Class + Sex, data = mydata)
summary(anova_2)

```

**Question (2 points):  Compare the two analyses.  What does the non-significant interaction term suggest about the relationship between L_RATIO and the factors CLASS and SEX?**

__Answer:__

It means the join effect of Class and Sex is not statistically higher than the sum of both effect individually.

### b.) 
__(2 points)__

+ For the model without CLASS:SEX (i.e. an interaction term), obtain multiple comparisons with the *TukeyHSD()* function. 
+ Interpret the results at the 95% confidence level (*TukeyHSD()* will adjust for unequal sample sizes). 

```{r Part_2b}


```

**Question (2 points):  first, interpret the trend in coefficients across age classes. What is this indicating about L_RATIO?  Second, do these results suggest male and female abalones can be combined into a single category labeled as 'adults?' If not, why not?**

***Answer: (Enter your answer here.)***

-----

(3)(a) (2 points) Use *combineLevels()* from the 'rockchalk' package to combine "M" and "F" into a new level, "ADULT". This will necessitate defining a new variable, TYPE, in mydata which will have two levels:  "I" and "ADULT". Present side-by-side histograms of VOLUME. One should display infant volumes and, the other, adult volumes. 

```{r Part_3a}


```

**Question (2 points): Compare the histograms.  How do the distributions differ? Are there going to be any difficulties separating infants from adults based on VOLUME?**

***Answer: (Enter your answer here.)***

(3)(b) (3 points) Create a scatterplot of SHUCK versus VOLUME and a scatterplot of their base ten logarithms, labeling the variables as L_SHUCK and L_VOLUME. Please be aware the variables, L_SHUCK and L_VOLUME, present the data as orders of magnitude (i.e. VOLUME = 100 = 10^2 becomes L_VOLUME = 2). Use color to differentiate CLASS in the plots. Repeat using color to differentiate by TYPE. 

```{r Part_3b}


```

**Question (3 points):  Compare the two scatterplots. What effect(s) does log-transformation appear to have on the variability present in the plot?  What are the implications for linear regression analysis? Where do the various CLASS levels appear in the plots? Where do the levels of TYPE appear in the plots?**

***Answer: (Enter your answer here.)***

-----

(4)(a) (3 points) Since abalone growth slows after class A3, infants in classes A4 and A5 are considered mature and candidates for harvest. Reclassify the infants in classes A4 and A5 as ADULTS. This reclassification can be achieved using *combineLevels()*, but only on the abalones in classes A4 and A5.  You will use this recoded TYPE variable, in which the infants in A4 and A5 are reclassified as ADULTS, for the remainder of this data analysis assignment.

Regress L_SHUCK as the dependent variable on L_VOLUME, CLASS and TYPE (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2 and Black Section 14.2). Use the multiple regression model: L_SHUCK ~ L_VOLUME + CLASS + TYPE. Apply *summary()* to the model object to produce results.

```{r Part_4a}


```

**Question (2 points):  Interpret the trend in CLASS levelcoefficient estimates? (Hint:  this question is not asking if the estimates are statistically significant. It is asking for an interpretation of the pattern in these coefficients, and how this pattern relates to the earlier displays).**

***Answer: (Enter your answer here.)***

**Question (2 points):  Is TYPE an important predictor in this regression? (Hint:  This question is not asking if TYPE is statistically significant, but rather how it compares to the other independent variables in terms of its contribution to predictions of L_SHUCK for harvesting decisions.)  Explain your conclusion.**

***Answer: (Enter your answer here.)***

-----

The next two analysis steps involve an analysis of the residuals resulting from the regression model in (4)(a) (Kabacoff Section 8.2.4, p. 178-186, the Data Analysis Video #2).

-----

(5)(a) (3 points) If "model" is the regression object, use model$residuals and construct a histogram and QQ plot. Compute the skewness and kurtosis. Be aware that with 'rockchalk,' the kurtosis value has 3.0 subtracted from it which differs from the 'moments' package. 

```{r Part_5a}


```

(5)(b) (3 points) Plot the residuals versus L_VOLUME, coloring the data points by CLASS and, a second time, coloring the data points by TYPE. Keep in mind the y-axis and x-axis may be disproportionate which will amplify the variability in the residuals. Present boxplots of the residuals differentiated by CLASS and TYPE (These four plots can be conveniently presented on one page using *par(mfrow..)* or *grid.arrange()*. Test the homogeneity of variance of the residuals across classes using *bartlett.test()* (Kabacoff Section 9.3.2, p. 222).  

```{r Part_5b}


```

**Question (3 points):  What is revealed by the displays and calculations in (5)(a) and (5)(b)? Does the model 'fit'?  Does this analysis indicate that L_VOLUME, and ultimately VOLUME, might be useful for harvesting decisions? Discuss.**  

***Answer: (Enter your answer here.)***

-----

There is a tradeoff faced in managing abalone harvest. The infant population must be protected since it represents future harvests. On the other hand, the harvest should be designed to be efficient with a yield to justify the effort. This assignment will use VOLUME to form binary decision rules to guide harvesting. If VOLUME is below a "cutoff" (i.e. a specified volume), that individual will not be harvested. If above, it will be harvested. Different rules are possible.

The next steps in the assignment will require consideration of the proportions of infants and adults harvested at different cutoffs. For this, similar "for-loops" will be used to compute the harvest proportions. These loops must use the same values for the constants min.v and delta and use the same statement "for(k in 1:10000)."  Otherwise, the resulting infant and adult proportions cannot be directly compared and plotted as requested. Note the example code supplied below.

-----

(6)(a) (2 points) A series of volumes covering the range from minimum to maximum abalone volume will be used in a "for loop" to determine how the harvest proportions change as the "cutoff" changes. Code for doing this is provided.

```{r Part_6a}

#idxi <- mydata$TYPE == "I"
#idxa <- mydata$TYPE == "ADULT"

#max.v <- max(mydata$VOLUME)
#min.v <- min(mydata$VOLUME)
#delta <- (max.v - min.v)/10000
#prop.infants <- numeric(10000)
#prop.adults <- numeric(10000)
#volume.value <- numeric(10000)

#total.infants <- sum(idxi)  
#total.adults <- sum(idxa)

#for (k in 1:10000) { 
#	value <- min.v + k*delta
#	volume.value[k] <- value
#	prop.infants[k] <- sum(mydata$VOLUME[idxi] <= value)/total.infants
#	prop.adults[k] <-  sum(mydata$VOLUME[idxa] <= value)/total.adults
#}

# prop.infants shows the impact of increasing the volume cutoff for
# harvesting. The following code shows how to "split" the population at
# a 50% harvest of infants.

#n.infants <- sum(prop.infants <= 0.5)
#split.infants <- min.v + (n.infants + 0.5)*delta  # This estimates the desired volume.
#split.infants

#n.adults <- sum(prop.adults <= 0.5)
#split.adults <- min.v + (n.adults + 0.5)*delta
#split.adults

```

(6)(b) (2 points) Present a plot showing the infant proportions and the adult proportions versus volume.value. Compute the 50% "split" volume.value for each and show on the plot.   

```{r Part_6b}


```

**Question (2 points):  The two 50% "split" values serve a descriptive purpose illustrating the difference between the populations. What do these values suggest regarding possible cutoffs for harvesting?** 

***Answer: (Enter your answer here.)***

-----

This part will address the determination of a volume.value corresponding to the observed maximum difference in harvest percentages of adults and infants. To calculate this result, the vectors of proportions from item (6) must be used. These proportions must be converted from "not harvested" to "harvested" proportions by using (1 - prop.infants) for infants, and (1 - prop.adults) for adults. The reason the proportion for infants drops sooner than adults is that infants are maturing and becoming adults with larger volumes.

-----

(7)(a) (1 point) Evaluate a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value. Compare to the 50% "split" points determined in (6)(a). There is considerable variability present in the peak area of this plot. The observed "peak" difference may not be the best representation of the data. One solution is to smooth the data to determine a more representative estimate of the maximum difference.

```{r Part_7a}


```

(7)(b) (1 point) Since curve smoothing is not studied in this course, code is supplied below. Execute the following code to create a smoothed curve to append to the plot in (a). The procedure is to individually smooth (1-prop.adults) and (1-prop.infants) before determining an estimate of the maximum difference. 

```{r Part_7b}

#y.loess.a <- loess(1 - prop.adults ~ volume.value, span = 0.25,
#	family = c("symmetric"))
# y.loess.i <- loess(1 - prop.infants ~ volume.value, span = 0.25,
#	family = c("symmetric"))
#smooth.difference <- predict(y.loess.a) - predict(y.loess.i)

```

(7)(c) (3 points) Present a plot of the difference ((1 - prop.adults) - (1 - prop.infants)) versus volume.value with the variable smooth.difference superimposed. Determine the volume.value corresponding to the maximum smoothed difference (Hint:  use *which.max()*). Show the estimated peak location corresponding to the cutoff determined.

```{r Part_7c}


```

(7)(d) (1 point) What separate harvest proportions for infants and adults would result if this cutoff is used? Show the separate harvest proportions (NOTE:  the adult harvest proportion is the "true positive rate" and the infant harvest proportion is the "false positive rate").

Code for calculating the adult harvest proportion is provided.

```{r Part_7d}

# (1 - prop.adults)[which.max(smooth.difference)]  # [1] 0.7416332

```

-----

There are alternative ways to determine cutoffs. Two such cutoffs are described below.

-----

(8)(a) (2 points) Harvesting of infants in CLASS "A1" must be minimized. The smallest volume.value cutoff that produces a zero harvest of infants from CLASS "A1" may be used as a baseline for comparison with larger cutoffs. Any smaller cutoff would result in harvesting infants from CLASS "A1."  

Compute this cutoff, and the proportions of infants and adults with VOLUME exceeding this cutoff. Code for determining this cutoff is provided. Show these proportions.

```{r Part_8a}

#volume.value[volume.value > max(mydata[mydata$CLASS == "A1" &
#  mydata$TYPE == "I", "VOLUME"])][1] # [1] 206.786

```

(8)(b) (2 points) Another cutoff is one for which the proportion of adults not harvested equals the proportion of infants harvested. This cutoff would equate these rates; effectively, our two errors:  'missed' adults and wrongly-harvested infants. This leaves for discussion which is the greater loss:  a larger proportion of adults not harvested or infants harvested?  This cutoff is 237.7383. Calculate the separate harvest proportions for infants and adults using this cutoff. Show these proportions.  Code for determining this cutoff is provided.  

```{r Part_8b}

# volume.value[which.min(abs(prop.adults - (1-prop.infants)))] # [1] 237.6391

```

-----

(9)(a) (6 points) Construct an ROC curve by plotting (1 - prop.adults) versus (1 - prop.infants). Each point which appears corresponds to a particular volume.value. Show the location of the cutoffs determined in (7) and (8) on this plot and label each. 

```{r Part_9}


```

(9)(b) (1 point) Numerically integrate the area under the ROC curve and report your result. This is most easily done with the *auc()* function from the "flux" package.   Areas-under-curve, or AUCs, greater than 0.8 are taken to indicate good discrimination potential. 

```{r Part_9b}


```

-----

(10)(a) (3 points) Prepare a table showing each cutoff along with the following:
 	1) true positive rate (1-prop.adults,
 	2) false positive rate (1-prop.infants),
 	3) harvest proportion of the total population
 	
```{r Part_10} 	


```
 	
**Question: (1 point) Based on the ROC curve, it is evident a wide range of possible "cutoffs" exist. Compare and discuss the three cutoffs determined in this assignment.**   

***Answer: (Enter your answer here.)***

-----

**Question (8 points):  Assume you are expected to make a presentation of your analysis to the investigators How would you do so?  Consider the following in your answer:**

1. Would you make a specific recommendation or outline various choices and tradeoffs?
2. What qualifications or limitations would you present regarding your analysis?
3. If it is necessary to proceed based on the current analysis, what suggestions would you have for implementation of a cutoff?  4)  What suggestions would you have for planning future abalone studies of this type? 

***Answer: (Enter your answer here.)***