---
title: "Lesson 08: Statistical Inference Estimation for Single Populations"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    df_print: paged
---

```{r knitr-setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(eval = TRUE)

opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 8) # for html
opts_knit$set(progress = FALSE, verbose = TRUE)

```


```{r additional-libraries, echo=FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

theme_set(theme_light())

# Theme Overrides
theme_update(plot.title = element_text(hjust = 0.5),
			 axis.text.x = element_text(size = 10),
			 axis.text.y = element_text(size = 10),
			 axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
			 legend.position = "top", legend.title = element_blank())

```

# Estimation for Single Populations

## 1.) 

> Assume a random sample of size 100 is drawn from a normal distribution with variance 1. The average
value of the sample is 50. Find a 95% confidence interval for the mean.

```{r echo=T}

```

## 2.)

> Assume the standard deviation for a normal distribution is equal to 100 units. Also assume we want
to estimate the unknown mean with a 95% confidence interval of total width 8 units. Calculate the
sample size required.


```{r echo=T}

```

## 3.) 

> A random sample of 1600 registered voters are contacted and asked a variety of questions. For one
question, 60% of the voters expressed approval and 40% disapproval. Calculate a 95% confidence
interval for the proportion expressing approval.


```{r echo=T}

```

## 4.)

> A random sample of consumers are presented with two beverages in random order and asked which
they prefer most. All the consumers expressed a preference. One beverage was preferred 85% of the
time. Use this number to determine how large a sample of consumers would be needed to generate a
95% confidence interval with an overall width just less than 2% (i.e. from 84% to 86%)?

```{r echo=T}

```

# Data Analysis

> Results of a laboratory analysis of calories and sodium content of major hot dog brands.
Researchers for Consumer Reports analyzed three types of hot dog: beef, poultry, and meat (mostly pork
and beef, but up to 15% poultry meat). Fifty four observations are reported.

## 1.)

> Create boxplots and find 95% confidence intervals for the mean amount of calories in each Type of hot
dog: beef, meat and poultry. Construct 99% one-sided lower confidence intervals for the mean amount
of calories in each Type of hot dog: beef, meat and poultry.


```{r echo=T}

```

## 2.)

> Find a 95% confidence interval for the variance in the amount of calories found for each type of hotdog:
beef, meat and poultry.


```{r echo=T}

```
